### 2020-12-09 Thesis-Chapter2-Background.tex
Work on wicked vs tame descriptions
Address more todos
Restructure contents of background chapter

% IDEA: create timeline https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline

@@

% IDEA: Create an appendix to go into greater detail?
% Could go into these, as well as the commented-out ones below
%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
% It would be nice to really understand these and how they're used today. committee_decrypting_2018 describes them.

@@

% % NOTE: based on the amendment's modifications, it is no longer about encryption. This has become more moderation and
% % CSAM focused, and the LAED Act has taken over the encryption side. A wise split (not that either bill is necessarily
% % good.)
% % https://www.engadget.com/earn-it-act-amendments-pass-senate-judiciary-committee-165030518.html

@@

% Prof. Walsh:
% Due to the dynamic nature of future judicial rulings that may expand or retract the application of exceptional access
% it is beyond the scope of this research to comment further...

@@

% IDEA: put this in an appendix too? Or its own chapter? Too much detail for the background section.
%
% For more, read donahue_comparative_2018 (Israel and China) and \cite{budish_encryption_2018} (framework for analysis)
%
% Since the enactment of the Regulation of Investigatory Powers Act 2000, the U.K. government had the ability to serve
% written notices requiring mandated decryption. Law enforcement and intelligence agencies also have explicit permission
% to perform ``computer network exploitation,'' and approach this thesis refers to as ``\ac{lawful-hacking}.'' 2016's
% Investigatory Powers Act expanded these capabilities and, among other additions, in section 253 created ``technical
% capability notices'' that the Secretary of State can issue to telecommunications operators mandating \ac{EA}
% \cite{legislature_2016}.

@@

% TODO: it is very important to know who these requirements apply to. Traditional telecoms, obviously. Facebook and
% Google? Signal and WhatsApp? Telecoms, not platforms or apps. This is the big differentiator. This can be expanded
% upon when I talk a bit about pen registers, etc.

@@

% FIXME: Engelsma: Consider moving this right after "encryption basics", as I have referred to many of these approaces
% already.

% There are several types of approaches:
% IDEA: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
% properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
% - Allow strong encryption (anyone with the users' private keys can see it)
% - Disallow encryption (anyone can see it)
% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
% - Require use of backdoor-ed algorithms, security via obscurity (anyone with the secret knowledge can see it)
% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
%   - Store extra in govt DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
%   - Split the key and store it multiple places
%   - See denning_taxonomy_1996
% - Use some alternative path
%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking

% FIXME: Separate out into separate, short chapter, probably after the wicked problems chapter. Also adds natural place
% to introduce my more "advanced" DFD syntax.

% TODO: include a more descriptive key to understand argument maps.

% - Prof. Dulimarta: Could you apply fuzzy logic to the argument map? Quantify it? -- I think this would run into limits
%   once you start comparing unquantifiable things, but it could be used to suss out the valid arguments, and would
%   probably be really interesting otherwise.

% Kalafut: I'm not sure that 2.5 and 2.6 really fit well as subsections of 2. Consider separating them into their own
% (short) chapter.
% Kalafut: Argument maps are used in chapter 4. I'm not sure that you need to mention argdown at all. If you do, I would
% only do so in the context of stating that it is the tool you are using to create argument maps.

\section{Argument Maps}
\label{sec-arg-maps-intro}

The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create \acp{argmap}.
\Acp{argmap} are graphical representations of the logical structures and relationships between statements, premises, and
conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
\acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some
software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.

Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo}
shows the basic structure of arguments. Statements are positions on issues. Some statements are based on values that
cannot be addressed through argument. Statements and arguments may support or attack one another, and arguments support
or undercut other arguments.

\begin{figure}[ht]
    \centering\CaptionFontSize
    \myincludeargument{demo}
    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
    \label{fig-args-demo}
\end{figure}

\section{Threat Models and Data Flow Diagrams}
\label{sec-threat-model-intro}

Threat modeling is an important step in analyzing the security at the systems level. Using models abstracts away fine
details and focuses on the architecture, processes, and dataflows in a system. Models assist in the understanding of
current systems and the prevention of problems in new systems, both of which are important when facing the prospect of
designing an \ac{EA} mechanism built on top of a complex and aging technology stack.

Threat modeling begins with the questions, ``What are you building?'' and ``What can go wrong?''
\cite{shostack_threat_2014}. Answering the first question accurately is crucial, particularly for the field of
cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. Answering the second
question depends on the types of attackers to worry about, and determines the scope of threats to be considered. ``What
are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go wrong?''
cannot be faithfully answered without knowing what is being built.

\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of
primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
\acp{DFD} in this style to analyze the threats already present and those that \ac{EA} would introduce.

\begin{figure}[ht]
    \centering\CaptionFontSize
    \includegraphics[width=\linewidth]{dfds/build/DFD-demo.png}
    \caption[\Acs{DFD} Demo]{A demonstrative example of a \acf{DFD}.}
    \label{fig-dfd-demo}
\end{figure}

>>>>

% Could create timeline https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline

% TODO: Does this read well here? Used to be after Regulatory Environment section.

% There are several types of approaches:
%  My selections here are also rather sparse. Create another appendix? Possibly create table of
% properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
% - Allow strong encryption (anyone with the users' private keys can see it)
% - Disallow encryption (anyone can see it)
% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
% - Require use of backdoor-ed algorithms, security via obscurity (anyone with the secret knowledge can see it)
% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
%   - Store extra in govt DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
%   - Split the key and store it multiple places
%   - See denning_taxonomy_1996
% - Use some alternative path
%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking

@@

@@

% % Note: based on the amendment's modifications, it is no longer about encryption. This has become more moderation and
% % CSAM focused, and the LAED Act has taken over the encryption side. A wise split (not that either bill is necessarily
% % good.)
% % https://www.engadget.com/earn-it-act-amendments-pass-senate-judiciary-committee-165030518.html

@@

@@

% Put this in an appendix too? Or its own chapter? Too much detail for the background section.
%
% For more, read donahue_comparative_2018 (Israel and China) and \cite{budish_encryption_2018} (framework for analysis)
%
% Since the enactment of the Regulation of Investigatory Powers Act 2000, the U.K. government had the ability to serve
% written notices requiring mandated decryption. Law enforcement and intelligence agencies also have explicit permission
% to perform ``computer network exploitation,'' and approach this thesis refers to as ``\ac{lawful-hacking}.'' 2016's
% Investigatory Powers Act expanded these capabilities and, among other additions, in section 253 created ``technical
% capability notices'' that the Secretary of State can issue to telecommunications operators mandating \ac{EA}
% \cite{legislature_2016}.

@@

% It is very important to know who these requirements apply to. This is for telecoms, not Facebook, Google, Signal, or
% WhatsApp. That's a little unclear, but the vagueness is a necessity, since I don't actually know this very well.
% Consider removing the internation section altogether because it's complicated and I'm not sure I'm adding a whole lot
% of value.

@@

### 2020-12-09 Thesis-Chapter6-ThreatModel.tex
Restructure contents of background chapter

This chapter consists of a threat model and analysis of a specific device key escrow \ac{EA} proposal, Stefan Savage's
``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}.
Lacking a concise name, this proposal will be referred to by its acronym, \ldawmsR. This chapter assumes that the
proposal is accompanied by some legal framework regulating its application and use, though it is not the focus of this
analysis, and nothing is assumed about it in particular. Note that the creators of the proposal intentionally chose the
phrase ``lawful access'' over ``\acl{EA}'' in order to emphasize the role of legal process and dispel the negative
stigma that accompanies \ac{EA}. I refer to \ldawmsr as an \ac{EA} proposal, but use the term entirely in the spirit of
their paper.

% TODO: remove this paragraph if I end up changing EA to LA

% TODO: Dulimarta: only looking at iPhone DAR? -- So I have to make clear that we are limiting scope here

@@

\modelitem{Protection of civil liberties}{The system should include measures to prevent mass surveillance and abuse by
political regimes that violate civil rights. This includes technologically enforcing the fundamental right to privacy
but upon warranted search pursuant to the rule of law.}

@@

% FIXME: add current hacking techniques.

@@

\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes that the legal process used to obtain the
warrant is valid. Additionally, all forms of surveillance, even when protected from mass surveillance, require
oversight. Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA}
proposals. However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}

@@

Answering ``what are we building'' and comparing it with the state of the art requires an understanding of the state of
the art. Apple iPhones are among the most secure consumer mobile phones, and serve as the point of reference for
\ldawmsR. \myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to their
public documentation \cite{apple_2020}.

@@

\begin{table}[h]
  \caption{Central Concepts in the \ldawmsr Proposal}
  \label{table-ldawmsr-concepts}
  \FlushLeft
  \small
  \begin{tabular}{ |l|p{6.4cm}|p{6.4cm}| }
    \hline
    \thead{Concept} & \thead{Implementation} & \thead{Outcome} \\ \hline
    Self-escrow
    & The Class key is escrowed in encrypted form in the device's Secure Enclave (or equivalent) using a write-only
      component; i.e., software can update the key, but only an unexposed hardware interface can read it.
    & Since the Class key itself is being stored instead of the encryption scheme being changed, there is no change to
      the security of the underlying cryptographic protocols. Physical possession and partial disassembly is required,
      which introduces practical barriers to mass surveillance.
    \\ \hline
    Time vaulting
    & The hardware interface responds to requests only after proof of sustained possession for a ``lockup period (e.g.,
      72 hours).''
    & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for mass surveillance.
    \\ \hline
    Authorization
    & After the lockup period, the requestor must provide evidence of authorization. This comes from a shared secret
      between the device and the manufacturer which can be obtained via legal process.
    & Physical possession is now not enough; law enforcement (or attackers) must provide this secret information to the
      device to unlock it.
    \\ \hline
    Transparency
    & In addition to the device disassembly, the escrow agent modifies the device (e.g., burns a fuse) when it has been
      unlocked. The device firmware detects this modification.
    & The device user has evidence that the device has been unlocked, preventing covert usage.
    \\ \hline
  \end{tabular}
\end{table}

Device key escrow is considered the most promising technical \ac{EA} direction because of the unique capabilities that
secure hardware offers, and \ldawmsr demonstrates this capabilities perfectly. The greatest of these capabilities are
physical possession requirements and strong transparency. Physical possession requirements strongly mitigate mass
surveillance and, as Savage puts it, ``provides a more intuitive compatibility with common under-standings of the
government’s reasonable law enforcement powers (e.g., the ability to seize physical property under court order) than
more information-centric approaches, which may appear covert by comparison'' \cite{savage_lawful_2018}. Transparency is
also rooted in the physicality of the approach, wherein the device cannot be unlocked without making irreversible and
detectable changes, mitigating both mass surveillance and common criminal abuse.

@@

To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{HSM}, \acf{DID}, \acf{DSK}, and
\acf{DOS}.

% TODO: Check list of acronyms up to date here ^^

@@

\threattablebegin{Denial of Service}{dos}{Service Denied}
  D1 & Unlock artifacts recovery
  & \ii{S6} could be used to inject corrupt data to the DID-DSK database, potentially overwriting actual unlock artifact
    data.
  & Verify integrity of new data before entering it into the database, and disallow overwriting entirely.
\threattableend

% TODO: Platform abuser destroys the unlock artifacts database through physical attack

@@

\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, and \ii{E1} \ii{E2} are each low risk either because they
fail to produce useful results, they can be well mitigated, or they are a threat that is already present in
state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and mitigate the
risk posed by these threats as part of the design process, but they do not represent fatal flaws to a design. Except in
combination with other threats, they are not worthy more concern here.

@@

It is important to give \ldawmsr (and similar \ac{DAR} \ac{EA} proposals) its due. The worst case scenario still
precludes mass surveillance and includes non-repudiation that access occurred. This is a significant gain for \ac{EA}
proposals in terms of transparency and protection of civil liberties.

@@

\paragraph*{Protection of civil liberties} This goal was defined as defending against mass surveillance and abuse by
authoritarian regimes. This proposal does an excellent job of protecting against mass surveillance due to its
physicality. Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian
regimes. Prior to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign
countries. After a leak, these regimes could get into any device they want. If one lived in a country that was willing
to take and destroy their devices to read their contents without legal due process, this would be an unwelcome
technology.

>>>>

This chapter consists of a threat model and analysis of a specific device key escrow \ac{EA} proposal, Stefan Savage's
``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}. The
proposal is specific to \acl{DAR} for mobile devices. Lacking a concise name, this proposal will be referred to by its
acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by some legal framework regulating its
application and use, though it is not the focus of this analysis, and nothing is assumed about it in particular. Note
that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
\ac{EA} proposal, but use the term entirely in the spirit of their paper.

@@

\modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
political regimes that violate civil rights. This includes technologically enforcing the fundamental right to privacy
but upon warranted search pursuant to the rule of law.}

@@

@@

\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes that the legal process used to obtain the
warrant is valid. Additionally, all forms of surveillance, even when protected from \ac{masssurv}, require oversight.
Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA} proposals.
However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}

@@

Answering ``what are we building'' and comparing it with the state of the art requires an understanding of the state of
the art. Apple iPhones are among the most secure consumer mobile phones, and serve as the point of reference for
\ldawmsR. As this analysis is limited to a specific proposal, it is also limited to this reference architecture.
\myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to Apple's public
documentation \cite{apple_2020}.

@@

\begin{table}[h]
  \caption{Central Concepts in the \ldawmsr Proposal}
  \label{table-ldawmsr-concepts}
  \FlushLeft
  \small
  \begin{tabular}{ |l|p{6.4cm}|p{6.4cm}| }
    \hline
    \thead{Concept} & \thead{Implementation} & \thead{Outcome} \\ \hline
    Self-escrow
    & The Class key is escrowed in encrypted form in the device's Secure Enclave (or equivalent) using a write-only
      component; i.e., software can update the key, but only an unexposed hardware interface can read it.
    & Since the Class key itself is being stored instead of the encryption scheme being changed, there is no change to
      the security of the underlying cryptographic protocols. Physical possession and partial disassembly is required,
      which introduces practical barriers to \ac{masssurv}.
    \\ \hline
    Time vaulting
    & The hardware interface responds to requests only after proof of sustained possession for a ``lockup period (e.g.,
      72 hours).''
    & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for \ac{masssurv}.
    \\ \hline
    Authorization
    & After the lockup period, the requestor must provide evidence of authorization. This comes from a shared secret
      between the device and the manufacturer which can be obtained via legal process.
    & Physical possession is now not enough; law enforcement (or attackers) must provide this secret information to the
      device to unlock it.
    \\ \hline
    Transparency
    & In addition to the device disassembly, the escrow agent modifies the device (e.g., burns a fuse) when it has been
      unlocked. The device firmware detects this modification.
    & The device user has evidence that the device has been unlocked, preventing covert usage.
    \\ \hline
  \end{tabular}
\end{table}

Device key escrow is considered the most promising technical \ac{EA} direction because of the unique capabilities that
secure hardware offers, and \ldawmsr demonstrates this capabilities perfectly. The greatest of these capabilities are
physical possession requirements and strong transparency. Physical possession requirements strongly mitigate mass
surveillance and, as Savage puts it, ``provides a more intuitive compatibility with common under-standings of the
government’s reasonable law enforcement powers (e.g., the ability to seize physical property under court order) than
more information-centric approaches, which may appear covert by comparison'' \cite{savage_lawful_2018}. Transparency is
also rooted in the physicality of the approach, wherein the device cannot be unlocked without making irreversible and
detectable changes, mitigating both \ac{masssurv} and common criminal abuse.

@@

To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{HSM}, \acf{DID}, \acf{DSK}, \acf{SN},
and \acf{DOS}.

@@

\threattablebegin{Denial of Service}{dos}{Service Denied}
  D1 & Unlock artifacts recovery
  & \ii{S6} could be used to inject corrupt data to the DID-DSK database, potentially overwriting actual unlock artifact
    data.
  & Verify integrity of new data before entering it into the database, and disallow overwriting entirely.
  \\ \hline
  D2 & Unlock artifacts recovery
  & Determined adversaries could physically destroy the medium storing the DID-DSK database.
  & The \ac{HSM} should be in a physically secure and guarded facility, and the storage medium should be stable and
    persistent.
\threattableend

@@

\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low risk either
because they fail to produce useful results, they can be reasonably mitigated, or they are a threat that is already
present in state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and
mitigate the risk posed by these threats as part of the design process, but they do not represent fatal flaws to a
design. Except in combination with other threats, they are not worthy more concern here.

@@

It is important to give \ldawmsr (and similar \ac{DAR} \ac{EA} proposals) its due. The worst case scenario still
precludes \ac{masssurv} and includes non-repudiation that access occurred. This is a significant gain for \ac{EA}
proposals in terms of transparency and protection of civil liberties.

@@

\paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
authoritarian regimes. This proposal does an excellent job of protecting against \ac{masssurv} due to its physicality.
Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior
to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After
a leak, these regimes could get into any device they want. If one lived in a country that was willing to take and
destroy their devices to read their contents without legal due process, this would be an unwelcome technology.
